---
title: "Testing for generality of pollinator recognition in *Heliconia*: Analysis of aviary data"
author: "D.G. Gannon, A.S. Hadley, U.G. Kormann, F.A. Jones, M.G. Betts"
output:
  github_document:
    pandoc_args: --webtex
bibliography: /Users/dusty/Documents/zotero_library.bib
---

```{r setup, message=FALSE, warning=FALSE}

library(tidyverse)
library(here)

dist_mat <- function(X){
  G <- X%*%t(X)
  ones <- rep(1, nrow(X))
  d <- diag(G)
  D <- d%*%t(ones) - 2*G + ones%*%t(d)
  return(D)
}

euc_dist <- function(v1, v2){
  sq_diff <- (v1-v2)^2
  
  return(sqrt(sum(sq_diff)))
}

```

Here we detail the calculations for hummingbird bill and *Heliconia* flower trait matching. 

### Hummingbird bill morphology

For hummingbird bill shape, we photographed wild-caught hummingbirds' bills against 0.25-inch gridded paper before releasing them. We then used `ImageJ` to measure the length of the bill along the outer (top) surface of the bill using a the line segment feature. To measure curvature, we utilized the methods of @temeles2009 in which we related the outer edge of the bill to a circle. The curvature of a circle is measured as the inverse of its radius. To estimate the radius of a circle with the same curvature as a bird's bill, we connected two points on the curved section of the bill, effectively drawing a chord on the circle. We then measured the angle between the chord and the line tangent to the bill at one end of the chord, known as the *angle of declension*. Straight-forward trigonomentric identities then yield $\hat{r} = \frac{c/2}{\sin{\theta}}$, where $c$ is the length of the chord (mm) and $\theta$ is the angle of declension (radians). Thus, we use $\hat{r}^{-1}$ as an estimate of the curvature of a bill or flower.


```{r}
# Load bird data

  bills <- read_csv(
    file = here("Data", "bird_bill_morph.csv")
  )

# compute curvature for bird bills
  bills <- bills %>% mutate(
    bill_chordtan_angle_rad = bill_chordtan_angle_deg*(pi/180)
  )
  
  bills <- bills %>% mutate(
    bill_curve = ((bill_chordL_mm/2)/(sin(bill_chordtan_angle_rad)))^(-1)
  )

# add column for the radius as well  
  bills <- bills %>% mutate(
    bill_r = (bill_chordL_mm/2)/(sin(bill_chordtan_angle_rad))
  )


  

```

### Plant morphology

```{r}

  flowers <- read_csv(
    file = here("Data", "heliconia_morphology.csv")
  ) 

# subset for focal species
  flowers <- flowers[
    which(flowers$Species %in%
            c("H.hirsuta", "H.rostrata",
              "H.tortuosa", "H.wagneriana")),
  ]

# compute curvature
  flowers <- flowers %>% mutate(
    angle_rad = angle*(pi/180)
  )
  
  flowers <- flowers %>% mutate(
    curvature = ((chord_length/2)/(sin(angle_rad)))^(-1)
  )
  
# compute radius as well
  flowers <- flowers %>% mutate(
    radius = (chord_length/2)/(sin(angle_rad))
  )
  

```


### Trait mismatch

We compute the average degree of mismatch as the distance between the means for each species in the $\mathbb{R}^2$ euclidean space with one axis as length (mm) and the other the radius of the arc of the curve (mm).

```{r}

  birds <- c("RTAH","GREH")
  plants <- c("H.hirsuta", "H.rostrata", "H.tortuosa", "H.wagneriana")
  
  mismatch <- as.data.frame(matrix(nrow = 8, ncol = 5))
  names(mismatch) <- c("bird", "plant", "mean_dist",
                       "sd_dist", "n")
  
  for(i in 1:length(birds)){
    for(j in 1:length(plants)){
      bird_i <- subset(bills, Species==birds[i])
      plant_i <- subset(flowers, Species==plants[j])
      nbirds <- nrow(bird_i)
      nflowers <- nrow(plant_i)
      N <- nbirds + nflowers
      
      # create matrix of bill measurements
      mat <- as.matrix(bird_i[,which(
        names(bird_i) %in% c("bill_length_mm", "bill_r")
      )])
      
      # bind matrix with flower measurements
      mat <- rbind(
        mat, as.matrix(
          plant_i[, which(
            names(flowers) %in% c("total_length", "radius")
          )]
        )
      )
      
      dist_ij <- dist_mat(mat)
      dists <- sqrt(as.double(dist_ij[(N-nflowers+1):N, 1:nbirds]))
      mismatch[(length(plants)*(i-1)+j), ] <-
        c(birds[i], plants[j], mean(dists), 
          sd(dists), length(dists))
    }
  }


```


### Some stats for the MS

```{r}

  rtah <- subset(bills, Species=="RTAH")

  mean(rtah$bill_length_mm)
  sd(rtah$bill_length_mm)
  
  mean(rtah$bill_curve)
  sd(rtah$bill_curve)

```

```{r}

 flwr_summary <- flowers[,
   which(names(flowers) %in% 
           c("Species", "total_length", "curvature"))
 ] %>% 
  group_by(Species) %>%
  summarise(., meanL=mean(total_length),
            s_L = sd(total_length),
            meanK=mean(curvature),
            s_K=sd(curvature))

  flwr_summary

```








